---
title: "Practice with t-tests"
output: html_document
date: "2022-10-27"
---

## Reviewing the t-test

Let's conduct a 2-sample t-test to compare body fat percentage between men and women.

The data set presents the body fat percentage for a sample of men and women who all work out at the same gym. In general, men are leaner than women, with a body fat percentage centering around 15% whereas for women it tends to center around 20%. Let’s see if the same is true of these data.

## Instructions

Use the information you learned from reading chapter 5 and doing homework 21 to conduct a t-test comparing mean body fat percentage between the male and female groups. But don’t just rush to the test – follow the recommended analysis workflow:
```{r}
rm(list=ls())
library(tidyverse)
library(here)

fat<- read.csv(here("data", "body_fat_percentage.csv"))

```
gender as a factor 
```{r}
fat$gender<- as.factor(fat$gender)
```

1.  Plot the data and look at expected relationships
```{r}
ggplot(fat, aes(x= gender, y = body_fat_percent))+
  geom_boxplot( fill = "black", color= "lightgray")+
  geom_point(aes(x = gender , y = body_fat_percent), color = "blue", alpha = 0.5, position = position_jitter(width = 0.2)) +
  labs(
    title = "gender vs. bodyfat",
    x = "gender",
    y = "body_fat_%"
  )
 theme_minimal()
  
```

2.  Calculate the mean and standard deviation for the data and examine the resulting output.
```{r}
fatcalc <- fat %>% 
  group_by(gender) %>% 
  summarise(
    meanfat = mean(body_fat_percent),
    sdfat = sd(body_fat_percent))
View(fatcalc)
```

3.  Discuss with your partner:
    -   Review in the book: What are the assumptions about the data appropriate for a t-test? When you examine the plots and summary data, do they appear to meet those assumptions?
    -   Based on examination of the plots and summary data, do you expect that there is a statistically significant difference in mean body fat percentage between males and females in this sample?
4.  Conduct a t-test and then interpret the results. What do they tell you about the null vs. alternative hypotheses?
```{r}
t_test_result <- t.test(body_fat_percent ~ gender, data = fat)


print(t_test_result)
```

