---
title: "21 Oct 2024 - review"
format: html
editor: visual
---

Download the dataset called "birds_and_mammals.csv" from today's module and put it in your Data folder.

Set up your workspace and read in the .csv. Name it "critters"

```{r}
rm(list=ls())
library(tidyverse)
library(here)
animals<- read.csv(here("data", "birds_and_mammals.csv"))
```

Filter just for mammals in the order Carnivora and make a new dataset called "carnivores"

```{r}
carnivores <- animals %>%
  filter(Order == "Carnivora")
```

create a histogram of maximum longevity for carnivores.

```{r}
ggplot(carnivores, aes(x = max_longevity_years)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Maximum Longevity for Carnivores",
       x = "Maximum Longevity (years)",
       y = "Frequency") +
  theme_minimal()
```

Now make a scatterplot comparing max_longevity_years vs. adult_wt_g

```{r}
ggplot(carnivores, aes(x = adult_wt_g, y = max_longevity_years)) +
  geom_point(color = "green", alpha = 0.6) +
  labs(title = "Scatterplot of Maximum Longevity vs. Adult Weight",
       x = "Adult Weight (grams)",
       y = "Maximum Longevity (years)") +
  theme_minimal()
```

Use code (rather than scrolling the data) to determine which species it is that has the highest maximum longevity?

```{r}
longest_lived_species <- animals %>%
  filter(max_longevity_years == max(max_longevity_years, na.rm = TRUE)) %>%
  select(Species, max_longevity_years)
```

Use code to determine the two species between 2.5e+05 and 3e+05 in adult body mass.

```{r}
species_wt_range <- animals %>%
  filter(adult_wt_g >= 2.5e+05 & adult_wt_g <= 3e+05) %>%
  select(Species, adult_wt_g)
```

Now make a boxplot of max longevity as a function of family. Set the y-axis label to "maximum longevity, years". Make the points red and somewhat transparent.

```{r}
ggplot(animals, aes +
  geom_boxplot(color = "red", alpha = 0.5, width = 0.2) +  
  labs(title = "Boxplot of Maximum Longevity by Family",
       x = "Family",
       y = "Maximum Longevity, Years") +
  theme_minimal()
```
